---
title: "Project1"
output: html_document
---

S04 Var01, Var02
S05 Var 02

```{r}
all_set <- read.csv("https://raw.githubusercontent.com/lincarrieli/Data624/main/Project1_Group2.csv", header = TRUE)

s04 <- subset(all_set, group == "S04")
s05 <- subset(all_set, group == "S05")
```

```{r}
s04Var01 <- s04$Var01
s04Var02 <- s04$Var02
s05Var02 <- s05$Var02
```

Initial visualization for S04 Var02, Var03 and S04 Var 02
```{r}
s04 %>%
  ggplot(aes(x=SeriesInd, y=Var01)) + geom_line() + ggtitle("s04Var01")

s04 %>%
  ggplot(aes(x=SeriesInd, y=Var02)) + geom_line() + ggtitle("s04Var02")

s05 %>%
  ggplot(aes(x=SeriesInd, y=Var02)) + geom_line() + ggtitle("s05Var02")
```


```{r}
library(tseries)
library(tidyverse)
library(fpp2)
library(readxl)
s04Var01.ts <- ts(s04Var01, frequency = 365)
s04Var01.ts1 <- na.remove(s04Var01.ts)

s04Var01.ts1 %>%
  decompose(type="multiplicative")%>%
  autoplot() + xlab("Time") +
  ggtitle("Classical multiplicative decomposition of s04Var01")
```
```{r}
s04Var02.ts <- ts(s04Var02, frequency = 365)
s04Var02.ts %>%
  decompose(type="multiplicative") %>%
  autoplot() + xlab("Time") +
  ggtitle("Classical multiplicative decomposition of s04Var02")
```


```{r}
s05Var02.ts <- ts(s05Var02, frequency = 365)

s05Var02.ts1 <- na.remove(s05Var02.ts)
s05Var02.ts1 %>%
  decompose(type="multiplicative") %>%
  autoplot() + xlab("Time") +
  ggtitle("Classical multiplicative decomposition of s04Var02")
```

Each variable seems to display distinct characteristics. s04Var02 has no apparent trend but do exhibit seasonality based on the decomposition plot; it also has a few outliers. s04Var03 has a upward trend and starting to go downward with clear seasonality. s05Var02 shows a downward trend at the beginning of the series and seemed to remine flat before starting to go upward; three extreme outliers can be observed from the plot.

Examine the nature of gaps/missing values. Using the approach of computing the =average squared difference across gaps

```{r}
SeriesInd.ts <- ts(s04$SeriesInd)

gaps <- diff(SeriesInd.ts) > 1

gaps <- c(FALSE, gaps)
gaps.df <- data.frame("SeriesInd" = s04$SeriesInd, "AfterGap" = gaps)

gaps.df <- gaps.df %>%
  mutate("s04Var01" = s04$Var01, "s04Var01.diff" = s04$Var01 - lag(s04$Var01))


sqdiff_across_gaps_s04Var01 <- gaps.df %>%
  filter(AfterGap) %>%
  filter(s04Var01.diff > -50) %>%
  select(s04Var01.diff)

sqdiff_across_gaps_s04Var01 <- sqdiff_across_gaps_s04Var01^2
sqdiff_across_gaps_s04Var01 <-
  mean(sqdiff_across_gaps_s04Var01$s04Var01.diff)
sqdiff_across_gaps_s04Var01
```

Computing the average squared difference between successive entries for s04Var01:

```{r}
sqdiff_across_all <- gaps.df %>%
  filter(s04Var01.diff < 50) %>%
  select(s04Var01.diff)

sqdiff_across_all <- sqdiff_across_all^2
sqdiff_across_all <- mean(sqdiff_across_all$s04Var01.diff)
sqdiff_across_all
```

The small difference in values of mean squared difference between gaps and successive values are small enough to suggest a pause in data generating process, and not missing data.


```{r}
s04Var01.ts <-ts(s04Var01)

autoplot(s04Var01.ts) +
  xlab("Day") +
  ylab("Value") +
  ggtitle("Daily Value of s04Var01")

s04Var02.ts <- s04 %>%
  filter(Var02 < 15000000) %>%
  select(Var02) %>%
  ts()

autoplot(s04Var02.ts) +
  xlab("Day") +
  ylab("Value") +
  ggtitle("Daily Value of s04Var02")

s05Var02.ts <- s05 %>%
  filter(Var02 < 15000000) %>%
  select(Var02) %>%
  ts()

autoplot(s05Var02.ts) +
  xlab("Day") +
  ylab("Value") +
  ggtitle("Daily Value of s05Var02")
```

```{r}
library(forecast)
library(ggplot2)
s04Var01_rwf <- rwf(s04Var01.ts, h = 140)
s04Var02_rwf <- rwf(s04Var02.ts, h = 140)
s05Var02_rwf <- rwf(s05Var02.ts, h = 140)

s04Var01_drwf <- rwf(s04Var01.ts, h = 140, drift = TRUE)
s04Var02_drwf <- rwf(s04Var02.ts, h = 140, drift = TRUE)
s05Var02_drwf <- rwf(s04Var02.ts, h = 140, drift = TRUE)

#s04Var01_mean <- mean(s04Var01.ts, h = 140)
#s04Var02_mean <- mean(s04Var02.ts, h = 140)
#s05Var02_mean <- mean(s05Var02.ts, h = 140)

autoplot(s04Var01.ts) +
  autolayer(s04Var01_rwf, series = "Naive", PI = FALSE) +
  autolayer(s04Var01_drwf, series = "Drift", PI = FALSE) 
 
autoplot(s04Var02.ts) +
  autolayer(s04Var02_rwf, series = "Naive", PI = FALSE) +
  autolayer(s04Var02_drwf, series = "Drift", PI = FALSE)

autoplot(s05Var02.ts) +
  autolayer(s05Var02_rwf, series = "Naive", PI = FALSE) +
  autolayer(s05Var02_drwf, series = "Drift", PI = FALSE) 
  
  
```
```{r}
s04Var01_rmse_rwf_nodrift <- tsCV(s04Var01.ts, rwf, drift = FALSE, h = 1)
s04Var01_rmse_rwf_nodrift <- sqrt(mean(s04Var01_rmse_rwf_nodrift^2, na.rm = TRUE))

s04Var01_rmse_rwf_drift <- tsCV(s04Var01.ts, rwf, drift = TRUE, h = 1)
s04Var01_rmse_rwf_drift <- sqrt(mean(s04Var01_rmse_rwf_drift^2, na.rm = TRUE))

s04Var02_rmse_rwf_nodrift <- tsCV(s04Var02.ts, rwf, drift = FALSE, h = 1)
s04Var02_rmse_rwf_nodrift <- sqrt(mean(s04Var02_rmse_rwf_nodrift^2, na.rm = TRUE))

s04Var02_rmse_rwf_drift <- tsCV(s04Var02.ts, rwf, drift = TRUE, h = 1)
s04Var02_rmse_rwf_drift <- sqrt(mean(s04Var02_rmse_rwf_drift^2, na.rm = TRUE))

s05Var02_rmse_rwf_nodrift <- tsCV(s05Var02.ts, rwf, drift = FALSE, h = 1)
s05Var02_rmse_rwf_nodrift <- sqrt(mean(s05Var02_rmse_rwf_nodrift^2, na.rm = TRUE))

s05Var02_rmse_rwf_drift <- tsCV(s05Var02.ts, rwf, drift = TRUE, h = 1)
s05Var02_rmse_rwf_drift <- sqrt(mean(s05Var02_rmse_rwf_drift^2, na.rm = TRUE))
```

```{r}
s04Var01_rmse_rwf_nodrift
s04Var01_rmse_rwf_drift

s04Var02_rmse_rwf_nodrift
s04Var02_rmse_rwf_drift 

s05Var02_rmse_rwf_nodrift
s05Var02_rmse_rwf_drift
```
For all models, the better performing model is random walk with no drift.

Exponential soomthing is suitable for data with clear trend or seasonality, and Holt’s linear trend method allows the forecasting of data with a trend. I  apply apply both methods to all three variables and would expect Exponential smoothing would better fit s04Var02, and Holt's would fit better with s04Var01 and s05Var02.

```{r}
s04Var02_ses <- ses(s04Var02.ts, h = 140)
summary(s04Var02_ses)
```

```{r}
s05Var02_ses <- ses(s05Var02.ts, h = 140)
summary(s05Var02_ses)
```
The optimized simple exponential smoothing method computed $\alpha = 0.9999$, making this method almost indistinguishable from the random-walk forecast.

```{r}
s04Var02_ses <- ses(s04Var02.ts, h = 140)
summary(s04Var02_ses)
```

```{r}
s05Var02_ses <- ses(s05Var02.ts, h = 140)
summary(s05Var02_ses)
```
For s04Var02 and s05Var02, the $\alpha = 0.4$, indicating that some weight is given to observations from the more distant past.
```{r}
s04Var02_1.ts <- s04Var02.ts %>%
  na.remove()
s04Var02_ses <- ses(s04Var02_1.ts, h = 140)

autoplot(s04Var02_ses) +
  autolayer(fitted(s04Var02_ses), series="Fitted") +
  ylab("Value") + xlab("Day")
```
```{r}
s05Var02_1.ts <- s05Var02.ts %>%
  na.remove()
s05Var02_ses <- ses(s05Var02_1.ts, h = 140)

autoplot(s05Var02_ses) +
  autolayer(fitted(s05Var02_ses), series="Fitted") +
  ylab("Value") + xlab("Day")
```
```{r}
s04VAr01_hw <- holt(s04Var01.ts1,
         damped = TRUE, h=140)

autoplot(s04VAr01_hw) +
  autolayer(s04VAr01_hw, series="Hold", PI=FALSE)+
  guides(colour=guide_legend(title="Daily forecasts of s04Var01"))

s04VAr02_hw <- holt(s04Var02.ts,
         damped = TRUE, h=140)

autoplot(s04VAr02_hw) +
  autolayer(s04VAr02_hw, series="Hold", PI=FALSE)+
  guides(colour=guide_legend(title="Daily forecasts of s04Var02"))

s05VAr02_hw <- holt(s05Var02.ts,
         damped = TRUE, h=140)

autoplot(s05VAr02_hw) +
  autolayer(s05VAr02_hw, series="Hold", PI=FALSE)+
  guides(colour=guide_legend(title="Daily forecasts of s05Var02"))
```

